---
output:
  html_document: default
  pdf_document: default
---
Is Solar Energy A Political Issue?
========================================================

### Tuckman, Jonah






Research question:
--------------------------------------------------------

I would like to attempt to understand whether there is a correlation between the emerging field of solar energy and states political alignment.

Data:
--------------------------------------------------------

### Data source:


https://catalog.data.gov/dataset/energy-generation-by-state-and-technology-2009/resource/bb0da868-f498-4c47-b61f-da217897198f


### Data collection:

I pulled the data off of the government data website and then in an R file trimmed it to hold only the needs for my research. I then found election results and added them to my csv file.

### Cases:

The best case would be that there is a correlation between political alignment and solar energy output, ie a higher percentage of (most likely democratic) states have a positive solar output.

### Variables:

My two variables are:  
1. Solar Output: Although numerical, I am using it as a categorical, unordered variable as I am focusing on the binary question of yes solar output vs no solar output.  
2. Political Alignment: This is a categorical unordered variable as well in a binary sense which can be either R (Republican) or D (Democrat).

### Type of study:


This is an observational study as I am merely collecting and analyzing the data rather than running an experiment on a population.

### Scope of inference:

This study will not be able to make definitive statements about the population because it is an observational study. It will, however, be able to provide information about a correlational relationship rather than a causational relationship.
### Load data:
This sets the working directory and then imports my original downloaded csv dataset

```{r}


setwd('/Users/JonahTuckman/Desktop/Statistics/Statistics-MATH-240/Final Project')


statsData <- read.csv('State RE Generation-Table 1.csv') ## Original dataset with many extra energy types

```

### Data clean-up:

I also had to replace the commas in the set so they could be used as integers.

```{r}

neededCols <- c("State","Solar") ## Focusing on states and Solar output
statsData <- statsData[neededCols] ## Taking only those columns stated above
Voting <- read.csv("Voting.csv")
statsData <- merge(statsData, Voting, by="State")
write.csv(statsData, "SolarData.csv") ## Writing this new file


statsData <- read.csv("SolarData.csv") ## Reading this new file
statsData['X'] <- NULL ## Deleted unneeded column

replaceCommas<-function(X){
  X<-as.numeric(gsub("\\,", "", X))
}


statsData$Solar <- replaceCommas(statsData$Solar)

```

Exploratory data analysis:
--------------------------------------------------------



```{r}
RepData <- subset(statsData, statsData$Voting == 'R')
DemData <- subset(statsData, statsData$Voting == 'D')

demSolarNum <- subset(DemData, DemData$Solar>0)
demSolarPerc <- (nrow(demSolarNum) / nrow(DemData) * 100)
print(demSolarPerc)
## 28.57% of Democratic States have a positive solar output

RepSolarNum <- subset(RepData, RepData$Solar > 0)
RepSolarPerc <-(nrow(RepSolarNum) / nrow(RepData) * 100)
print(RepSolarPerc)
## 10% of Republican States have a positive solar output

### Show Data
countsRep <- table(RepData$Solar < 1, RepData$Solar > 1)
barplot(countsRep, main="Republican Solar Usage", horiz=TRUE,
        xlab = "Political Alignment", col = c("darkblue", "red"),
        legend = rownames(countsRep), xlim = c(0, 30), ylim = c(0, 3))

countsDem <- table(DemData$Solar < 1, DemData$Solar > 1)
barplot(countsDem, main="Democratic Solar Usage", horiz=TRUE,
        xlab = "Political Alignment", col = c("darkblue", "red"),
        legend = rownames(countsDem), xlim = c(0, 30), ylim = c(0, 3))



```





Data:
--------------------------------------------------------
A one page print out of my data is attached.

<!-- Do not forget to print out a page of your data and turn it in with your proposal-->
